Secure Composition \\ of Robust and Optimising Compilers

### 2024-05-27 csf2/main.tex
rw

  When the data independent timing mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory is related to a specification trace with \colorbox{lightgray}{no exposed secrets}.
\end{exampleenv}
\begin{figure}[!htb]
  $$
  % {
  \begin{array}{ccc}
    & \Lscct &        \\
    Active & \mid & Inactive \\\hline
    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} \\
    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} \\
    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} \\
    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} \\
    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
    \vdots & \mid & \vdots \\
    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} \\
    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} \\
    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} \\
    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} \\
  \end{array}
  % }
  $$
  $$
  % {
  \begin{array}{ccc}
          & Spec & \\
     Active & \mid & Inactive \\\hline
     \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
     \ev{Alloc}\ \loc_{y}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
     \ev{Use}\ \loc_{y}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\lock\\
     \ev{Use}\ \loc_{x}\ 1;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\lock\\
     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
     \vdots & \mid & \vdots \\
      \ev{Use}\ \loc_{x}\ 12;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\lock\\
      \emptyevent & \mid & \ev{Branch}\ 1;\lock\\
      \ev{Dealloc}\ \loc_{y};\unlock & \mid & \ev{Dealloc}\ \loc_{y};\unlock\\
      \ev{Use}\ \loc_{y}\ 6;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\lock\\
  \end{array}
  % }
  $$
  \caption{Traces for \Cref{ex:lscct}.}
  \label{fig:ex-cct}
\end{figure}

@@

This section discusses work on robust compilation (\Cref{subsec:relw:seccomprtp}) and on other secure compilation criteria (\Cref{subsec:relw:seccompcrit}).
Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms}, \gls*{cct}, and \gls*{ss}, this section then presents relevant related work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs,subsec:relw:ssmechs}).

\subsection{Secure Compilation as Robust Preservation}\label{subsec:relw:seccomprtp}

The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
No existing work is concerned with composing robustly safe compilers.
%Parts of these works consider languages with different trace models and our technical setup can handle this. 
The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) but use it in the context of protocols.
The work is missing the upper and lower compositions as well as the generality to support different trace models.
% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) as well as in an earlier version of this work~\cite{kruse2022csc}.
% However, they do not demonstrate the scalability of the approach by means of a case study.

\subsection{Other Secure Compilation Criteria}\label{subsec:relw:seccompcrit}

While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
Fully abstract compilation~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
It was shown~\cite{abate2021faandrc} that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.

\subsection{Memory Safety Mechanisms}\label{subsec:relw:msmechs}

@@

\subsection{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}

@@

\subsection{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}

With the advent of speculative execution in processors, it is possible to exploit this unusual paradigm identify secrets.
SPECTECTOR~\cite{guarnieri2018spectector} models speculative execution formally and the present work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from SPECTECTOR.
A subsequent line of work~\cite{fabian2022automatic} shows how to modularly define and compose formal semantics that models different speculative execution variants. 
Similar in spirit, recent work~\cite{fabian2024lift} demonstrates how to lift existing secure compilation proofs for security properties related to speculative execution to more powerful properties, i.e., properties defined for languages with a trace model that enables more vulnerabilities, such as combinations of different speculative execution exploits. 
To this end, they define a lifting framework and demonstrate how a secure compilation proof for a weaker property can be lifted to a stronger one. 
Contrary to the compositionality framework presented in this paper, these works are only concerned with composing semantics of the respective programming languages, while in this paper, the secure compilers themselves are composed.

>>>>

  When the data independent timing mode is off, the execution yields events in similar fashion to before (\Cref{subsec:ltms,subsec:lsms,subsec:lms}).
  But, if it is turned on, then the branching event \colorbox{yellow}{does not fire} anymore and both reading and writing to memory is related to a specification trace with \colorbox{lightgray}{no exposed secrets}.
\end{exampleenv}
\begin{figure}[!htb]
	\vspace{-1em}
  $$
  % {
  \begin{array}{ccc}
    & \Lscct &        \\
    Active & \mid & Inactive \\\hline
    \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{x}\ 12;\comp;\unlock} \\
    \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} & \mid & \obj{\ev{Alloc}\ \loc_{y}\ 12;\comp;\unlock} \\
    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
    \obj{{\ev{iGet}}\ \loc_{x}\ 0;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 0;\comp;\lock} \\
    \obj{{\ev{iSet}}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Set}\ \loc_{y}\ 0\ \mathtt{'H'};\comp;\lock} \\
    \obj{{\ev{iGet}}\ \loc_{x}\ 1;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 1;\comp;\lock} \\
    \text{\colorbox{yellow}{$\obj{\emptyevent}$}} & \mid & \obj{\ev{Branch}\ 0;\comp;\lock} \\
    \vdots & \mid & \vdots \\
    \obj{{\ev{iGet}}\ \loc_{x}\ 12;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{x}\ 12;\comp;\lock} \\
    \text{\colorbox{yellow}{$\obj{\emptyevent}$}}& \mid & \obj{\ev{Branch}\ 1;\comp;\lock} \\
    \obj{\ev{Dealloc}\ \loc_{y};\comp;\unlock} & \mid & \obj{\ev{Dealloc}\ \loc_{y};\comp;\lock} \\
    \obj{{\ev{iGet}}\ \loc_{y}\ 6;\comp;}\text{\colorbox{lightgray}{$\obj{\lock}$}} & \mid & \obj{\ev{Get}\ \loc_{y}\ 6;\comp;\lock} \\
  \end{array}
  % }
  $$
  $$
  % {
  \begin{array}{ccc}
          & Spec & \\
     Active & \mid & Inactive \\\hline
     \ev{Alloc}\ \loc_{x}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
     \ev{Alloc}\ \loc_{y}\ 12;\unlock & \mid & \ev{Alloc}\ \loc_{x}\ 12;\unlock\\
     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
     \ev{Use}\ \loc_{x}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 0;\lock\\
     \ev{Use}\ \loc_{y}\ 0;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 0;\lock\\
     \ev{Use}\ \loc_{x}\ 1;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 1;\lock\\
     \emptyevent & \mid & \ev{Branch}\ 0;\lock\\
     \vdots & \mid & \vdots \\
      \ev{Use}\ \loc_{x}\ 12;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{x}\ 12;\lock\\
      \emptyevent & \mid & \ev{Branch}\ 1;\lock\\
      \ev{Dealloc}\ \loc_{y};\unlock & \mid & \ev{Dealloc}\ \loc_{y};\unlock\\
      \ev{Use}\ \loc_{y}\ 6;\text{\colorbox{lightgray}{$\unlock$}} & \mid & \ev{Use}\ \loc_{y}\ 6;\lock\\
  \end{array}
  % }
  $$
  \caption{Traces for \Cref{ex:lscct}.}
  \label{fig:ex-cct}
\end{figure}

@@

This section discusses robust compilation, other secure compilation criteria and work related to the properties preserved in the case study.
% (\Cref{subsec:relw:seccomprtp}) 
% (\Cref{subsec:relw:seccompcrit}).
% Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls*{ms}, \gls*{cct}, and \gls*{ss}, 
% Then, this section discusses work related 
% relevant related work as well 
% (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs,subsec:relw:ssmechs}).

\paragraph*{Secure Compilation as Robust Preservation}\label{subsec:relw:seccomprtp}

The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
No existing work is concerned with composing robustly safe compilers.
%Parts of these works consider languages with different trace models and our technical setup can handle this. 
The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) but use it in the context of protocols.
The work does not consider the generality to support different trace models.
% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:sequential}) as well as in an earlier version of this work~\cite{kruse2022csc}.
% However, they do not demonstrate the scalability of the approach by means of a case study.

\paragraph*{Other Secure Compilation Criteria}\label{subsec:relw:seccompcrit}

While this paper focuses on the robust preservation framework~\cite{abate2019jour}, other secure compilation criteria exist.
The survey on formal approaches to secure compilation~\cite{patrignani2019survey} discusses a broad spectrum already, while this section presents a very high-level overview.
Fully abstract compilation~\cite{abadi1999fullabstraction} states that a compiler should preserve and reflect observational equivalence between source and target programs.
Abate \emph{et al.}~\cite{abate2021faandrc} showed that fully abstract compilers robustly preserve program properties that are either trivial or meaningless.
As a mitigation for this, the authors presented a categorical approach based on maps of distributive laws~\cite{watanabe2002modl}, which they call many maps of distributive laws.
Maps of distributive laws have been investigated before as a possible secure compilation criterion~\cite{tsampas2020catsc}.
Other approaches are extensions of the compiler correctness criterion as discussed in other work~\cite{patterson2019next700} or the introduction of opaque observations~\cite{vu2021reconciling} to reconcile compiler optimisations with security.
Note that this work also presents secure compilers that are optimising, but contrary to the other~\cite{vu2021reconciling}, provides a formal account of these in the robust preservation framework.
% Lastly, the authors of this paper have presented ongoing work~\cite{patrignani2023blame} on a weaker robust preservation criteria based on the concept blame.

\paragraph*{Memory Safety Mechanisms}\label{subsec:relw:msmechs}

@@

\paragraph*{Cryptographic Constant Time Mechanisms}\label{subsec:relw:cctmechs}

@@

\paragraph*{Speculation Safety Mechanisms}\label{subsec:relw:ssmechs}

With the advent of speculative execution in processors, it is possible to exploit this unusual paradigm leak secrets.
This work has taken inspiration in the sematics of $\Lspec$ and in \Cref{def:trace:ss} from recent formalisations~\cite{guarnieri2018spectector,fabian2022automatic}.
\MPin{
  we take SS from exorcising.
  we do not take SNI from spectector because it is a hypersafety.
  other work defined similar props.
}
% A subsequent line of work~\cite{fabian2022automatic} shows how to modularly define and compose formal semantics that models different speculative execution variants. 
% Recent work~\cite{fabian2024lift} demonstrates how to lift existing secure compilation proofs for security properties related to speculative execution to more powerful properties, i.e., properties defined for languages with a trace model that enables more vulnerabilities, such as combinations of different speculative execution exploits. 
% To this end, they define a lifting framework and demonstrate how a secure compilation proof for a weaker property can be lifted to a stronger one. 
% Contrary to the compositionality framework presented in this paper, these works are only concerned with composing semantics of the respective programming languages, while in this paper, the secure compilers themselves are composed.
